---
title: "MonopolyProject"
output: html_document
author: Donart Tota
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Solving Monopoly
If you have played monopoly, unlike me, you might thing that there is no cerain strategy when it comes to buying property in monopoly because it is all due to chance. But you might be wrong and in this project we will try to show which color property you should buy in order to win. 

##Basic functions

The following function will be used throughout the project. This function helps us create vectors that keep shifting down by one. We will use this to create almost all our matrices. 

```{r}
#Function that takes the the end of the vector and puts it infront
cycleVec <- function(vec, k){
  front <- tail(vec, k)               # take the end of the vector
  back <- head(vec, length(vec) - k)  # take head up to tail
  c(front, back)                      # flip them. 
}

```


##Step 1: Preliminary Frequencies
In this part we will create a simple model for the game not including any extra rules. Here we will include the rolling matrix, community chest, go to jail, and chance matrix. 

First we will add the rolling matrix. Here we are building the matrix based on the rolling of two die. 
```{r pressure, echo=FALSE}
# We create a roll matrix 40x40 since that is how many spots we have. The probabilities will be all the way from 0 to 6/36. We go from 2 moves all the way up to 12. 

rollMatrix <- matrix(0, nrow=40, ncol=40)
probs <- c(0,1,2,3,4,5,6,5,4,3,2,1,rep(0,28))/36
for(i in 0:40){
  rollMatrix[,i] <- cycleVec(probs, i)
}
colSums(rollMatrix)
```
As we see all the columns add up to 1 and the probabilities as we want. 

###Adding chance
Now we add a more difficult one which is chance. Chance has 16 cards and 10 of them move you in the following places:

Advance to Go
Go to Jail
Go to Illinois Avenue
Go to St. Charles
Take a walk on the Boardwalk (Go to Boardwalk)
Go back three spaces
Go to nearest Utility (depends on which Chance location you are on)
Go to nearest Railroad - there are two of these cards
Go to Reading Railroad

First we create an identity matrix
```{r}
#adding chance
chanceMatrix <- matrix(0, nrow=40, ncol=40)
probs <- c(rep(0,39),1)
for(i in 0:40){
  chanceMatrix[,i] <- cycleVec(probs, i)
}
```

Then we change the columns that are chance columns which are 8,23, and 37. 

Chances that we stay in the same place is 6/16. Chances that we move to nearest rail road is 2/16. Chances for all the moving cards is 1/16
```{r}
chanceMatrix[,8] <- rep(0,40)
chanceMatrix[c(1,11,25,12,40,6,5,13),8] <- 1/16
chanceMatrix[16,8] <- 2/16
chanceMatrix[8,8] <- 6/16

chanceMatrix[,23] <- rep(0,40)
chanceMatrix[c(1,11,25,12,40,6,20,29),23] <- 1/16
chanceMatrix[26,23] <- 2/16
chanceMatrix[23,23] <- 6/16

chanceMatrix[,37] <- rep(0,40)
chanceMatrix[c(1,11,25,12,40,34,13),37] <- 1/16
chanceMatrix[6,37] <- 3/16
chanceMatrix[37,37] <- 6/16

colSums(chanceMatrix)
```

As we see we get 1 in the columns now we do community chest. 

###Adding community chest
Community chest has 16 cards and two of them move you around. So probability to sta is 14/16, and probability to go to jail, or advance to go is 1/16. 

Community chest is located at spots 3, 18 and 34 

```{r}
#Ading community chest identity matrix
communityMatrix <- matrix(0, nrow=40, ncol=40)
probs <- c(rep(0,39),1)
for(i in 0:40){
  communityMatrix[,i] <- cycleVec(probs, i)
}

communityMatrix[,c(3,18,34)] <- rep(0,40)
communityMatrix[c(1,11),3] <- 1/16
communityMatrix[c(1,11),18] <- 1/16
communityMatrix[c(1,11),34] <- 1/16

communityMatrix[3,3] <- 14/16
communityMatrix[18,18] <- 14/16
communityMatrix[34,34] <- 14/16

colSums(communityMatrix)
```
All columns sum up to 1 so all is good. 

###Adding jail
Now we add the easiest one which is go to jail. Here we create an identity matrix and we just change row 11,31 to 1. 
```{r}
#Ading jail
goToJailMatrix <- matrix(0, nrow=40, ncol=40)
probs <- c(rep(0,39),1)
for(i in 0:40){
  goToJailMatrix[,i] <- cycleVec(probs, i)
}

goToJailMatrix[31,31] = 0
goToJailMatrix[11,31] = 1

colSums(goToJailMatrix)

```

###Calculating the steady state vectors
Now we just multiply all the matrices together and we get the steady state vector. Here me mutiply community matrix before chance matrix since one can send you to the other. 
```{r}
P = communityMatrix %*% chanceMatrix %*% goToJailMatrix %*% rollMatrix

eigenStuff <- eigen(P)
eigenVecs <- eigenStuff$vectors
steadyStateVec <- eigenVecs[,1]
steadyStateVec <- steadyStateVec/sum(steadyStateVec)
as.numeric(steadyStateVec[c(11, 25, 1,20,26,6,16,19,21,22,29,17,24,12)])
```
As we see the steady state vector matches the table seen in the paper. We should notice that in my code 1 is Go and in the paper 40 is Go. So everything in the index counter is shifted by 1. 


##Part 2 Adding stay in jail
Apparently there is a rule that allows you to stay in jail as much as you want. In order to simulate this we will add two columns that will represent 2 of the three turns you can stay in jail. Those columns will be in the matrix that is the same as the roll matrix but has two new columns 41, 42. 41 sends you to 42, and 42 sends you to 11

```{r}
# We copy those matrices. 
stayInJailLonger <- rollMatrix

# Add two columns of 0s
stayInJailLonger <- cbind(stayInJailLonger, c(rep(0,40)), c(rep(0,40)))

# Add two rows of 0s
stayInJailLonger <- rbind(stayInJailLonger, c(rep(0,42)), c(rep(0,42)))

# You can go from 41 to 42, and 42 to 11, 5/6 times unless you get doubles. 
stayInJailLonger[42,41] <- 5/6
stayInJailLonger[11,42] <- 5/6
```



Where we can go with doubles. Basically 11 + 2... +12. Everthing is 1/36 since that is the chance of getting a double. We do this for each column 
```{r}
stayInJailLonger[c(13,15,17,19,21,23),41] <- 1/36
stayInJailLonger[c(13,15,17,19,21,23),42] <- 1/36

colSums(stayInJailLonger)
```
All are 1s so we should be doing alright. 


###Adding jail

As usual this is the easiest part. It will send us to state 41 if we fall in state 31. 

```{r}
goToJailMatrix2 <- matrix(0, nrow=42, ncol=42)
probs <- c(rep(0,41),1)
for(i in 0:42){
  goToJailMatrix2[,i] <- cycleVec(probs, i)
}

# Send the player to the first part of the jail. 
goToJailMatrix2[41,31] <- 1 
goToJailMatrix2[31,31] <- 0
colSums(goToJailMatrix2)
```

###Ading Chance

In this part we will add chance which is a little anoying. It is litteraly the same as above the only difference is that now we go to column 11 as the first state of the matrix. I could have probably just changed the previous one after copying it but this seemed easier. 

We create the identity matrix 42x42 and then we change the probabilities. 

```{r}
chanceMatrix2 <- matrix(0, nrow=42, ncol=42)
probs <- c(rep(0,41),1)
for(i in 0:42){
  chanceMatrix2[,i] <- cycleVec(probs, i)
}

chanceMatrix2[c(1,12,25,5,13,6,40,41,6),8] <- 1/16
chanceMatrix2[16,8] <- 2/16
chanceMatrix2[8,8] <- 6/16

chanceMatrix2[c(1,12,25,20,26,29,6,40,41,6),23] <- 1/16
chanceMatrix2[26,23] <- 2/16
chanceMatrix2[23,23] <- 6/16


chanceMatrix2[c(1,12,25,34,13,6,40,41),37] <- 1/16
chanceMatrix2[6,37] <- 3/16
chanceMatrix2[37,37] <- 6/16

colSums(chanceMatrix2)
```
All 1s so we should be doing ok. 


### Adding Comunity Chest
As above next thing to add is community chest. In this part the only thing that is changing is where we are going to jail and the size of the matrix which is 42x42. 

We change the two states that we can move to which is go and jail 1. the other probability is that we stay in the same place. 
```{r}
communityMatrix2 <- matrix(0, nrow=42, ncol=42)
probs <- c(rep(0,41),1)
for(i in 0:42){
  communityMatrix2[,i] <- cycleVec(probs, i)
}

communityMatrix2[c(1,41),3] <- 1/16
communityMatrix2[3,3] <- 14/16

communityMatrix2[c(1,41),18] <- 1/16
communityMatrix2[18,18] <- 14/16

communityMatrix2[c(1,41),34] <- 1/16
communityMatrix2[34,34] <- 14/16

colSums(communityMatrix2)
```
As we see we get one. 

###Transition matrix

Now we build the transition matrix. The order of multiplying doesn't change. Now the only thing is that we are using the new 42x42 matrices. 
```{r}
P2 = communityMatrix2 %*% chanceMatrix2 %*% goToJailMatrix2 %*% stayInJailLonger

eigenStuff <- eigen(P2)
eigenVecs <- eigenStuff$vectors
steadyStateVec <- eigenVecs[,1]
steadyStateVec <- steadyStateVec/sum(steadyStateVec)
sum(steadyStateVec[c(11,41,42)])
steadyStateVec[c(25,1,26,6,21,20,19,17,29,16,13,22,12)]
```
The values are the same as professor Richeys. The only difference that I noticed is that he has rounded up to 4 significant digits. We get the same exact one for the jail

##Part 3 Adding doubles 

Apparently there is another rule in monopoly that if you role three times in a row doubles you will go to jail. Chances for that to happen are 1/216. So we will create a matrix that is similar to the identity matrix but it has 215/216 where the ones should be and it has 1/216 to send you to state one of jail.
```{r}
# Matrix similar to the identity one. 
doubleJail <- matrix(0, nrow=42, ncol=42)
probs <- c(rep(0,41),215/216)
for(i in 0:42){
  doubleJail[,i] <- cycleVec(probs, i)
}


#sending every other state to jail with probability 1/216
doubleJail[41,c(1:40)]=1/216
doubleJail[31,31]=0
doubleJail[41,31]=1

doubleJail[42,42]=1
doubleJail[41,41]=1
colSums(doubleJail)
```
Looks
```{r}
P3 = communityMatrix2 %*% chanceMatrix2 %*%  doubleJail %*% goToJailMatrix2 %*% stayInJailLonger

eigenStuff <- eigen(P3)
eigenVecs <- eigenStuff$vectors
steadyStateVec <- eigenVecs[,1]
steadyStateVec <- steadyStateVec/sum(steadyStateVec)
sum(steadyStateVec[c(11,41,42)])
```
We are off by .06%. I think this might be because I don't understand what to do in the roll three time case if you are in prison. But since it is such a small difference it should still not affect our final result. 

P.S. I really wish I had played monopoly before. 

##Part 4 Analysis

In this part we have to do the analysis. We have done all the difficult work so we just have to do some adding in the percentages and we will see which color is better to buy. 

In the part bellow we are going to group all the colors. In the csv file there is a column call group and each color is preresented by a number. The code bellow reads the csv file into a data frame and then saves all the rows where group is 1,2,3...8 on different colors. 

```{r}
MonopolyData <- read.csv("MonopolyData.csv")
purple = MonopolyData[MonopolyData$Group=="1",]
lightblue = MonopolyData[MonopolyData$Group=="2",]
maroon = MonopolyData[MonopolyData$Group=="3",]
orange = MonopolyData[MonopolyData$Group=="4",]
red = MonopolyData[MonopolyData$Group=="5",]
yellow = MonopolyData[MonopolyData$Group=="6",]
green = MonopolyData[MonopolyData$Group=="7",]
darkblue = MonopolyData[MonopolyData$Group=="8",]
```

The functions bellow takes a group and finds the sum of the frequency for a certain group.

```{r}

#Function for group steady state
groupPercentage<-function(x){
  x=as.data.frame(x)
  b1=x$X
  b2=sum(steadyStateVec[b1])
  return(b2)
}

#we can the function for every color and save the results on a table. This way we can use the vector to build the final matrix in the end. 

grouptable=c(groupPercentage(purple),groupPercentage(lightblue),groupPercentage(maroon),groupPercentage(orange),groupPercentage(red),groupPercentage(yellow),groupPercentage(green),groupPercentage(darkblue))
grouptable=grouptable*100 

grouptable

```

Now we build a function that calculates rent. Here we just multiply the steady state vector for a group with rent
```{r}

groupRent<-function(x){ 
  x=as.data.frame(x)
  b1=x$X
  b2=sum(steadyStateVec[c(b1)]*x$Rent)
   return(b2)
}

#same idea as above with the vector
grouptable1=c(groupRent(purple),groupRent(lightblue),groupRent(maroon),groupRent(orange),groupRent(red),groupRent(yellow),groupRent(green),groupRent(darkblue))

grouptable1

```
And finaly we calculate the rolls to break even following the insturctions in the paper and then put all the groups together. 
```{r}
breakEven<-function(x){
  x=as.data.frame(x)
  c1=sum(x$Cost)+sum(x$Total)
  c2=c1/groupRent(x)
  return(c2)
}

grouptable2=c(breakEven(purple),breakEven(lightblue),breakEven(maroon),breakEven(orange),breakEven(red),breakEven(yellow),breakEven(green),breakEven(darkblue))
names(grouptable2)=c("purple","light_blue","maroon","orange","red","yellow","green","dark_blue")

```

Now we put the three columns together. The rows are names in the same order that we have called the grouping function. 

```{r}
m=matrix(c(grouptable,grouptable1,grouptable2),ncol=3)
rownames(m)=c("Purple","Light Blue","Maroon","Orange","Red","Yellow","Green","Dark Blue")
colnames(m)=c("Group Frequency","Group Value","Rolls to break even")
m
```

The results that we get are very similar to the ones that were produced by professor Richey. There is clearly an error that was carried at some point that I added double roll because that's when i started not getting the exact numbers as before. 

##Conclusion 
From these results we can say that the orange properties are the best ones to buy since they have a small number of rolls to break even and a high group frequency. So when playing it would be more advantagous for a player to buy property in that area in order to win. The next best one to buy would be the red properties. 
